---
title: "Regression-Jonas"
output: html_document
---
```{r}
library(tidyverse)
ev <- read_csv("/Users/jonasbrockmoller/Documents/GitHub/ElectricVehiclesEDA/Dataset/ElectricCarData_Clean_Me.csv")
ev
```
This is my first try of a Simple Linear Regression

1. Show variables in a scatter plot (independent variable X = BatteryPack_size, depended variables Y = Range in km)
```{r}
  ggplot(ev, aes(x=Battery_Pack_Kwh, y=Range_Km)) + geom_point()
```
2. lm()-Command
```{r}
  model <- lm(Range_Km ~ Battery_Pack_Kwh, data = ev)
```
3. geom_line()
```{r}
  ggplot(ev, aes(x=Battery_Pack_Kwh, y=Range_Km)) + geom_point() + geom_smooth(method='lm', formula= y~x)
```
4. summary(model)
```{r}
summary(model)
```

4.1 Coefficients and results

  The results show that:
  - P-Value is very smal (smaller than 0.05) -> Null-Hypothesis can be discarded
  - R-squared is at 82.87% -> This means 82% of the variance of the dependend variable (Y=range in km) can be explained by this model -> good value
  - 3.850 is the slope of the graph (Steigung des Graphen). This means per 1 kwh more battery size the car can go 3.85 km more
  - 86.752 is the value of x=0. Does not make much sense for this example but it means with 0 battery size a ar can go 86 km.
  
```{r}
 plot(model)
```
  
